---
title: "Timberland Project Report"
author: "Yijia Yu"
date: "18/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

## 1. Introduction

  Timberland investment instruments are often used as an alternative asset class in the public and private pension fund managers' portfolios. As a long-term investment, timberland assets have the benefits of low correlation to the stock market, portfolio diversification, and inflation hedging. Therefore, timberland assets not only make the portfolio more diversified but also reduce the volatility. A powerful component of timberland return drivers is biological growth, which is independent of all factors typically impacting other investments and makes timberland assets distinguishing from others.
  
  Since the COVID-19 was declared a pandemic on March 11th, 2020, the world experienced the most severe economic recession since the great depression. In the first two quarters of 2020, the annual rate of growth in Canadian real GDP has dropped nearly 20%. Beginning on February 2020, stock markets worldwide became extremely volatile, with all major Wall Street in- dices experiencing their worst drop since 1987 followed by a quick rebound. Timberland as a alternative real asset becomes increasing popular for some Canadian pension-fund managers, as this kind of illiquid asset can provide institutional investors with strong diversification benefits and stable valuations.


#### Timberland Industry in Canada
  The forestry industry is an important contributor to Canada’s economy, generating \$25.8 billion to Canada’s GDP in 2018 and directly employing about 1209,940 people across the country. The forestry industry contributed 8% export in 2018 and generated \$2.9 billion revenues for provincial, territorial and federal governments in 2017. Top export markets for Canadian forestry are the United States (66%) and China (15%) in 2018. Stronger Canadian lumber and plup demand stimulated by U.S. housing starts and other industry demand makes a promising perspective for the Canadian forestry markets.

  The majority of Canada’s forests are owned by the provinces in which they are located, with just 1.6 percent being owned federally. The provincial and territorial governments are responsible for managing forest conservation, with strict laws in place to shape forest practices. These stringent laws are particularly important when it comes to halting logging practices and encouraging reforestation, as extensive deforestation was carried out by European settlers in Canada throughout the 18th and 19th centuries.
  
  The wood product manufacturing sector is the most lucrative of the forestry industries, with a nominal GDP of approximately 11 billion Canadian dollars in 2017. Western Forest Products, a lumber company based in British Columbia, is one of the key players in the forestry industry with an annual revenue of around 1.2 billion Canadian dollars in 2018. As a result, British Columbia employs by far the most people in the forestry and logging industry, followed by Quebec.
  
  The United States is the main trading partner of Canada when it comes to forest resources, with Canadian exports worth around 24.24 billion Canadian dollars in 2017. Lumber, sawmill and millwork products are the most traded Canadian forest products, with exports worth just over 15 billion Canadian dollars in 2018 globally.

  In spite of today’s challenging economic environment, the long-term supply and demand fundamentals of timberland bode well for the patient investor who thinks in terms of decades and not quarters. A growing Canadian and global population with a rising standard of living will continue to increase demand on a finite supply of timberland.

#### About this study

  The main purpose of this study is to investigate long-term financial performance for timberland asset in a mixed-asset portfolio. The modern portfolio theory will be used to construct the optimal portfolios, and to analyze its diversification effects. In order to find the time-varying optimized allocations, time-series GARCH model will be used.

### Data used in this study

1. Private timberland market

Returns of private equity timberland investment performance is approximated by the NCREIF Timberland Index (NTI), which is a quarterly time series composite return measure of investment performance of a large pool of individual U.S. timber properties acquired in the private market for investment purposes only. 

2. Public timberland market

As of July 2019, there is four publicly traded REITs that specialize in timberland. Together, they own more than 17 million acres of timberland properties in the U.S., Canada, New Zealand and other locations. The four REITs compose the timber sector of the FTSE NAREIT All REIT index are listed below:

- Potlatch 
- CatchMark
- Rayonier 
- Weyerhaeuser

3. To examine the role of timberland assets in a mixed portfolio, assets including large-cap stocks, small-cap stocks, returns of private- and public-equaty real estate, treasury bonds, and treasury bills are considered in this study. 
Returns on those asset classes are approximated by the S&P 500 Index (SP500), Russell 2000 Index (RU2000), National Council of Real Estate Investment Fiduciaries Property Index (NPI) and National Association of Real Estate Investment Trust (NAREIT) 10-Year Treasury bonds yield, and 3-month treasury bills respectively. 

<!-- ## 2. Methods  -->

<!-- ## 2.1 Modern portfolio theory -->
<!--   The modern portfolio theory introduced by Markowitz (1952) establishes the foundation of portfolio optimization and asset allocation strategies. This theory targets to optimize the portfolio returns through balancing the weighted combination of asset classes in the portfolio while minimize the risk. Mathematically, M risky assets (i=1,2,...,m) with individual returns $R = (R_1, R_2,...,R_m)^T$. Investors construct the portfolios by adjusting the weights of assets $w = (w_1, w_2,...,w_m)^T$ with constraints of $\sum_iw_i=1$ and $w_i>0$ (no short selling). The mean and covariance of returns are $$E(R) = \alpha = (\alpha_1,...,\alpha_m)^T$$,  -->
<!-- $$Cov(R) = \Sigma = \begin{bmatrix} -->
<!-- \Sigma_{1,1} & ... & \Sigma_{1,m}\\ -->
<!-- \vdots &  & \vdots\\ -->
<!-- \Sigma_{m,1} & ... & \Sigma_{m,m}\\ -->
<!-- \end{bmatrix}$$  -->

<!-- Therefore, the portfolio expected return can be calculated as $E(R) =w'\alpha$, and variance as $var(R) = w'\Sigma w$. Let's then assume that the portfolio risk $F$ is a function $w$ and $E(r)$. The optimized portfolio under the modern portfolio theory is found from the solution of the optimization problem -->
<!-- $$MinF(w, r)\\subject\ to\ w'\alpha=\alpha_0\\ \ w'1_m = 1 ,w_i>0$$ -->
<!-- where $\alpha_0$ is the assigned target return. Using the modern portfolio theory, one can study the efficient frontiers (portfolios with the minimized risk measure given a target return). Also this theory can determine the optimal asset allocations either in single and multiple periods.  -->



<!-- ## 2.2 Portfolio optimzation framework -->
<!-- 1. The mean variance (M-V) portfolio -->

<!-- Under the mean variance framework, a mixed-asset portfolio can be optimized by minimizing the risk subject to a given target $\mu$:  -->
<!-- $$\underset{w}{Min}\ var(R)=\frac12w^T\sum w \\ s.t. w^T \alpha=\alpha_0\\ \sum_{i=1}^{n}w_i=1 ,w_i>0$$  -->
<!-- Solving the optimization problem by iteratively changing the weights of assets, the mean-variance frontier represents return and risk combination of all possible efficient portfolio.  -->

<!-- To solve this problem, we define the lagrangian:  -->
<!-- $$L(w, \lambda_1, \lambda_2) = \frac12 w'\Sigma w+\lambda_1(\alpha_0-w'\alpha)+\lambda_2(1-w'1_m)$$ -->
<!-- derive the first-order conditions -->
<!-- $$\frac{\partial L}{\partial w} = 0_m= \Sigma w - \lambda_1\alpha-\lambda_21_m$$ -->
<!-- $$\frac{\partial L}{\partial \lambda_1} = 0= \alpha_0-w'\alpha$$ -->
<!-- $$\frac{\partial L}{\partial \lambda_2} = 0= 1-w'1_m$$ -->
<!-- Solve $w$ in terms of $\lambda_1, \lambda_2$: -->
<!-- $w_0 =\lambda_1\Sigma^{-1}\alpha+\lambda_2\Sigma^{-1}1_m$ -->
<!-- Solve for $\lambda_1,\lambda_2$ by substituting for w: -->
<!-- $$\alpha_0 = w_0'\alpha = \lambda_1(\alpha'\Sigma^{-1}\alpha) + \lambda_2(\alpha'\Sigma^{-1}1_m)$$ -->
<!-- $$1 = w_0'1_m = \lambda_1(\alpha'\Sigma^{-1}\alpha) + \lambda_2(\alpha'\Sigma^{-1}1_m)$$ -->
<!-- so we have: -->
<!-- $$\begin{bmatrix}\alpha_0\\1\end{bmatrix}= \begin{bmatrix}a&b\\b&c\end{bmatrix}\begin{bmatrix}\lambda_1\\\lambda_2\end{bmatrix}$$ -->
<!-- $a = \alpha'\Sigma^{-1}\alpha$, $b= \alpha'\Sigma^{-1}1_m$, $c = 1_m'\Sigma^{-1}1_m$. -->

<!-- With the given values of $\lambda_1$ and $\lambda_2$, the solution portfolio has the minimum variance equal to $$w_0'\Sigma w_0 \\ =\lambda_1^2(\alpha'\Sigma^{-1}\alpha)+2\lambda_1\lambda_2(\alpha'\Sigma^{-1}1_m)+\lambda_2(1_m\Sigma^{-1}1_m)  -->
<!-- \\=\begin{bmatrix}\lambda_1\\\lambda_2\end{bmatrix}'\begin{bmatrix}a&b\\b&c\end{bmatrix}\begin{bmatrix}\lambda_1\\\lambda_2\end{bmatrix}$$ -->
<!-- substituting $$\begin{bmatrix}\lambda_1\\\lambda_2\end{bmatrix}= \begin{bmatrix}a&b\\b&c\end{bmatrix}^{-1}\begin{bmatrix}\alpha_0\\1\end{bmatrix}$$ -->
<!-- gives optimal portfolio variance: -->
<!-- $$\begin{bmatrix}\alpha_0\\1\end{bmatrix}'\begin{bmatrix}a&b\\b&c\end{bmatrix}^{-1}\begin{bmatrix}\alpha_0\\1\end{bmatrix} \\= \frac1{ac-b^2}(c\alpha^2_0-2b\alpha_0+a)$$ -->




<!-- 2. Mean Contional Value at Risk (M-CVaR) framework  -->

<!-- We assume the random vector R has a probabiliy density function denoted by p(r). For a fixed vector w, we compute CDF associated with that vector $w$.  -->
<!-- $$\Psi(w,\gamma)= \int_{f(w,r)\leq\gamma}p(r)dr $$  -->
<!-- VaR is defined as the maximum loss that will not be exceeded within a given time period at a specified confidence level $\alpha\in (0,1)$, a portfolio's (1-$\alpha$)% VaR can be compute by  -->
<!-- $$VaR_{\alpha}(w) = min\{\gamma\in\mathbf{R}:\Psi(w,\gamma\geq\alpha\}$$  -->
<!-- Then the $CVaR_\alpha $ associated with the portfolio is -->
<!-- $$CVaR_\alpha(w) = \frac{1}{1-\alpha} \int_{f\leq VaR_\alpha (w)} f(w,r)p(r)dr$$ -->
<!-- Therefore, the mean-CVaR portfolio optimization is a problem of: -->
<!-- $$\underset{w}{Min}\ CVaR_\alpha(w) \\ s.t. w^T \alpha=\alpha_0\\ \sum_{i=1}^{n}w_i=1 ,w_i>0$$  -->
<!-- Rockafeller and Uryasev (2000) defined the simpler function to optimize the problem.  -->
<!-- $$F_\alpha(w,\gamma):=\gamma+\frac{1}{1-\alpha}\int_{f(w,r)\geq\gamma}(f(w, r)-\gamma)p(r)dr$$. Alternatively, $$F_\alpha(w,\gamma) = \gamma + \frac1{1-\alpha}\int[f(w,r)-\gamma]^+p(r)dr$$,  -->
<!-- where $[x]^+ = max(x,0)$.  -->
<!-- So that $F_\alpha(w,\gamma)$ is a convex function of $\gamma$ ; $VaR_\alpha$ is the minimizer and $CVaR$ is the minimum value of the convex function. -->



<!-- ## 2.3 Asset return volatility -->

<!--   Generalized autoregressive conditional heteroskedasticity (GARCH) models are used to exam the time-varying conditional variances. To guarantee a postive sign for the conditional variance, we used exponential GARCH model in our case. Our model is computed by using the R package **rugarch** -->



## 3. Results
```{r message=FALSE, warning=FALSE, include=FALSE}
library("devtools")
load("timberland.Rdata")
library(tseries)
library(forecast)
library(dygraphs)
library(PerformanceAnalytics)
require(ggplot2)
require(GGally)
library(psych)
```

## 3.1 Descriptive analysis 
This is a interactively graphical representation of selected quarterly return indices from 1987 Q1 to 2020 Q1. Some clear pattern could be observed. First, NCREIF is much less volatile than REIT, SP500, RU2000. Second, REIT follows SP500 closely but NTI doesn't. For example, both REIT and SP500 suffered from a big loss during 2008 financial crisis, whereas NCREIF kept a positive returns until 2009/Q2, where they only got a slightly negative return value. Some people said it's because timberland market tends to lag the public equity market. But some also argues that it's because of the biological growth of trees wouldn't stop as the market went into depression. 
```{r}
dygraph(all.qtr*100, main = "Quarterly returns of analyzed assets", ylab = "Return%")%>% 
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3))%>%
  dyRangeSelector()%>%
  dyLegend(width = 500)%>%
  dyOptions(colors = RColorBrewer::brewer.pal(7, "Dark2"))

```

The unconditional correlations of the return series were studied by calculating the correlation coefficients and by plotting the pairwise scatter plots, as showed in the following graph. In this portfolio, the correlation of NCREIF with all financial assets are uniformly low, except for the 3 month Treasury bill (CA). The correlation between big cap (SP500) and small cap (RU2000) are highly significant, and they're almost linearly correlated. Unlike private timber market, the timber REITs shows a closely correlation with the stock market and 10 year treasury bonds. Returns on most asset classes are positively skewed. However, by using the Jarque-Bera test statistic, most of asset returns are not normally distributed.  
```{r, message=FALSE}
allxts = na.omit(all.qtr)
ggpairs(as.data.frame(allxts))+theme_bw()
```


Summary statistics of all asset classes from 1987 Q1 to 2010 Q1 are reported in the following table. 
On a quarterly basis, NTI (private timberland index) has the highest mean return of 2.71% and a moderate standard deviation of 3.78%. Timber REITs have a highest standard deviation but a relative lower mean of 2.06%. Treasury bills have both lowest return and standard deviation. 
```{r}
describe(allxts*100)[,-1]
```
Three risk measures are applied to individual asset, and the results are reported in the following table. 
```{r}
res = rbind(
apply(allxts, 2, sd),
VaR(allxts, p=.95, method="modified"),
ES(allxts, p=.95, method="modified")
)
row.names(res)[1] = "SD"
row.names(res)[3] = "CVaR"
res
```

```{r message=FALSE, warning=FALSE}

# stationarity
apply(allxts,2,adf.test)
apply(allxts,2,pp.test)
```

```{r}
op = par(mfrow=c(1,2))
#test stationarity and normality
for(i in 1:ncol(allxts)){
  #plot acf
  acf(na.omit(allxts[,i])^2,lag.max = 100,xlab = colnames(allxts)[i], ylab = 'ACF', main=' ')
  #plot pacf
  pacf(na.omit(allxts[,i])^2,lag.max = 100,xlab = colnames(allxts)[i], ylab = 'PACF', main=' ')
}

```


## 3.2 Portfolio Optimization

```{r message=FALSE, warning=FALSE}
library(RColorBrewer)
library(fPortfolio)
pacman::p_load(matrixcalc,knitr,dygraphs,ggthemes,highcharter,viridis,tibbletime,timetk,tidyquant,tidyverse,fPortfolio,xts)
```

```{r}
#0. Prepare data
returns.data <- na.omit(allxts)
returns.data %>%tk_tbl()%>%head()
returns_ts <- as.timeSeries(returns.data)
spec <- portfolioSpec()
setSolver(spec) <- "solveRquadprog"
setNFrontierPoints(spec) <-25
#1. calculating equal weight portfolio
nAssets <- ncol(returns_ts)
setWeights(spec)<-rep(1/nAssets, times = nAssets) 
constraints <- 'LongOnly'
portfolioConstraints(returns_ts, spec, constraints)

# calculate the properties of the portfolio
# Now let us display the results from the equal weights portfolio, the assignment of weights, and the attribution of returns and risk.
ewPortfolio <- feasiblePortfolio( returns_ts, spec, constraints)
print(ewPortfolio)
col <- divPalette(ncol(returns_ts), "RdBu")
weightsPie(ewPortfolio, radius = 0.7, col = col)
mtext(text = "Equally Weighted MV Portfolio", side = 3, line = 1.5,
        font = 2, cex = 0.7, adj = 0)
weightedReturnsPie(ewPortfolio, radius = 0.7, col = col)
mtext(text = "Equally Weighted MV Portfolio", side = 3, line = 1.5,
        font = 2, cex = 0.7, adj = 0)
covRiskBudgetsPie(ewPortfolio, radius = 0.7, col = col)
mtext(text = "Equally Weighted MV Portfolio", side = 3, line = 1.5,
        font = 2, cex = 0.7, adj = 0)


# COMPUTE A MINIMUM RISK EFFICIENT PORTFOLIO!
minriskSpec <- portfolioSpec()
targetReturn <- getTargetReturn(ewPortfolio@portfolio)["mean"]
setTargetReturn(minriskSpec) <- targetReturn
minriskPortfolio <- efficientPortfolio(
  data = returns_ts,
  spec = minriskSpec,
  constraints = "LongOnly")
print(minriskPortfolio)

col <- qualiPalette(ncol(returns_ts), "Dark2")
weightsPie(minriskPortfolio, radius = 0.7, col = col)
mtext(text = "Minimal Risk MV Portfolio", side = 3, line = 1.5,
        font = 2, cex = 0.7, adj = 0)
weightedReturnsPie(minriskPortfolio, radius = 0.7, col = col)
mtext(text = "Minimal Risk MV Portfolio", side = 3, line = 1.5,
        font = 2, cex = 0.7, adj = 0)
covRiskBudgetsPie(minriskPortfolio, radius = 0.7, col = col)
mtext(text = "Minimal Risk MV Portfolio", side = 3, line = 1.5,
        font = 2, cex = 0.7, adj = 0)

# Tangency Portfolio
tgSpec <- portfolioSpec()
setRiskFreeRate(tgSpec) <- 0
tgPortfolio <- tangencyPortfolio(
  data = returns_ts,
  spec = tgSpec,
  constraints = "LongOnly")
print(tgPortfolio)
col <- seqPalette(ncol(returns_ts), "BuPu")
weightsPie(tgPortfolio, box = FALSE, col = col)
mtext(text = "Tangency MV Portfolio", side = 3, line = 1.5,
        font = 2, cex = 0.7, adj = 0)
weightedReturnsPie(tgPortfolio, box = FALSE, col = col)
mtext(text = "Tangency MV Portfolio", side = 3, line = 1.5,
        font = 2, cex = 0.7, adj = 0)
covRiskBudgetsPie(tgPortfolio, box = FALSE, col = col)
mtext(text = "Tangency MV Portfolio", side = 3, line = 1.5,
        font = 2, cex = 0.7, adj = 0)
#Now we are ready to start the optmization and pass data, specification and constraints to the function portfolioFrontier(). We can examine the result using a standard print() command.

# compute the efficient frontier
setNFrontierPoints(spec) <-25
frontier <- portfolioFrontier(returns_ts, spec, constraints)
print(frontier)

# plot efficient frontier
tailoredFrontierPlot(object = frontier, mText = "MV Portfolio - LongOnly Constraints",
                     risk = "Cov")

# plot weights
weightsPlot(frontier, col=rainbow(dim(returns.data)[2]))
text <- "Mean-Variance Portfolio - Long Only Constraints"
mtext(text, side = 3, line = 3, font = 2, cex = 0.9)
weightedReturnsPlot(frontier)
covRiskBudgetsPlot(frontier)

set.seed(12)
frontierPlot(object = frontier, pch = 19, cex = 0.5)
monteCarloPoints(object = frontier, mcSteps = 5000, pch = 19,
                   cex = 1,col="#D53E4F")
frontierpts <- frontierPoints(object = frontier)
lines(frontierpts, col = "blue", lwd = 2)
```

GARCH model estimation, Backtesting the risk model and Forecasting
```{r}
#Ncreif
xts.nti <- allxts[, "NTI"]
xts.npi <- allxts[, "NPI"]
REITs <- allxts[, "Timber REITs"]
# ARIMA
# ncreif
model.arima = auto.arima(xts.nti , trace = F , ic = 'bic')
model.arima
checkresiduals(model.arima)
model.arima3 = auto.arima(xts.npi, stationary = T, trace = F , ic = 'bic')
model.arima3
checkresiduals(model.arima3)
# REIT
model.arima2 = auto.arima(REITs, trace = T , ic = 'bic')
checkresiduals(model.arima2)

# box test
Box.test(model.arima$residuals^2,lag = 12,type="Ljung-Box")
Box.test(model.arima2$residuals^2, lag = 12, type="Ljung-Box")
Box.test(model.arima3$residuals^2, lag = 12, type="Ljung-Box")

#colnames(allxts)
for(i in 1:ncol(allxts)){
  y = allxts[,i]
  res=Box.test((y - mean(y))^2, lag = 12,type="Ljung-Box")
  print(res$p.value)
}

# GARCH model
library("rugarch")
spec = ugarchspec(variance.model = list(model="sGARCH",garchOrder=c(1,1)),
                  mean.model = list(armaOrder=c(1,1)),
                  distribution.model = "std")
# Nelson's egarch model
egarch.spec = ugarchspec(variance.model=list(model="eGARCH",garchOrder=c(1,1)),
                         mean.model=list(armaOrder=c(0,0)),  
                         distribution.model="std")
garch.fit1 = ugarchfit(egarch.spec, xts.nti)
# simulation
set.seed(123)
tseq = seq(as.Date("1987/1/1"), as.Date("2020/6/1"), "weeks")
sim = ugarchsim(garch.fit1,n.sim=length(tseq), n.start=0, m.sim=1, startMethod="sample")
simseries = xts(sim@simulation$seriesSim, order.by = tseq)
auto.arima(simseries , trace = T , ic = 'bic')

garch.fit1 = ugarchfit(egarch.spec, xts.nti)
# simulation
set.seed(1234)
sim = ugarchsim(garch.fit1,n.sim=2000, n.start=0, m.sim=1, startMethod="sample")
simseries = as.data.frame(sim@simulation$seriesSim)
auto.arima(simseries, stationary = TRUE , trace = T , ic = 'bic')


plot(garch.fit1, which="all")

# backtesting model
garchroll1 <- ugarchroll(egarch.spec, data=simseries, n.start = 1000, refit.every = 100, refit.window = "moving",VaR.alpha = 0.01,solver="hybrid", fit.control = list(scale = 1))
report(garchroll1, type = "VaR",VaR.alpha = 0.01, conf.level = 0.99)

plot(garchroll1, which="all")
plot(garchroll1, which=4)
# Forecasting Risk and VaR
garchfcst <- ugarchforecast(garch.fit1, n.ahead = 12)
garchfcst
plot(garchfcst,which=1)
plot(garchfcst,which=3)

garch.fit2=ugarchfit(egarch.spec,data=xts.nti, solver="hybrid", out.sample=5)
garchfcst2<-ugarchforecast(garch.fit2, data = NULL, n.ahead = 10, n.roll = 5, external.forecasts = list(mregfor = NULL, vregfor = NULL))
garchfcst2
plot(garchfcst2,which=2)
plot(garchfcst2,which=4)
```
