---
title: "Timberland Project Report"
author: "Yijia Yu"
date: "6/5/2020"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Introduction

Timberland investment instruments are often used as an alternative asset class in the public and private pension fund maneger's portfolio. Unlike the traditional financial tools such as stocks, bonds, cash, timberland asset has the benefit of low correlation to the stock market and the tendency of rise over in a long run. Therefore, timberland assets not only makes the portfolio more diversified but also minimizes the volatility. A powerful component of timberland return drivers is biological growth, which is independent of all factors typically impacting other investments and makes timberland asset distinguishing from others.

#### Timberland Industry in Canada
The forestry industry is a key contributor to Canada’s economy, with real GDP growth recorded at 0.7 percent in 2017. This comes as no surprise, as there were around 308.94 million hectares of forest area recorded in Canada in 2019

The majority of Canada’s forests are owned by the provinces in which they are located, with just 1.6 percent being owned federally. The provincial and territorial governments are responsible for managing forest conservation, with strict laws in place to shape forest practices. These stringent laws are particularly important when it comes to halting logging practices and encouraging reforestation, as extensive deforestation was carried out by European settlers in Canada throughout the 18th and 19th centuries.
The wood product manufacturing sector is the most lucrative of the forestry industries, with a nominal GDP of approximately 11 billion Canadian dollars in 2017. Western Forest Products, a lumber company based in British Columbia, is one of the key players in the forestry industry with an annual revenue of around 1.2 billion Canadian dollars in 2018. As a result, British Columbia employs by far the most people in the forestry and logging industry, followed by Quebec.
The United States is the main trading partner of Canada when it comes to forest resources, with Canadian exports worth around 24.24 billion Canadian dollars in 2017. Lumber, sawmill and millwork products are the most traded Canadian forest products, with exports worth just over 15 billion Canadian dollars in 2018 globally.

In spite of today’s challenging economic environment, the long-term supply and demand fundamentals of timberland bode well for the patient investor who thinks in terms of decades and not quarters. A growing canadian and global population with a rising standard of living will continue to increase demand on a finite supply of timberland.

#### About this study

The main purpose of this study is to investigate long-term financial performance metrics of both private and public timberland investments. I will employ the modern portfolio theory to evaluate its
performance and to analyze its diversification effects. 


## 2. Methods 

## 2.1 Build Portfoilos
To examine the role of timberland assets in a mixed portfolio, four assets including large-cap stocks, small-cap stocks, treasury bonds, and treasury bills are considered in this study. 
Returns on those asset classes are proxied by the S&P 500 Index (SP500), Russell 2000 Index (RU2000), 10-Year Treasury bonds yield, and 3-month treasury bills. 

The NCREIF Timberland Index (NTI) is used to proxy the returns for private-equity timberland investments in the U.S and Canada. Moreover, **the NTI is the best available benchmark for U.S. timberland investments**. Quarterly
data from 1987Q1 to 2020Q1 are used in this study.

## 2.2 Asset return volatility

Generalized autoregressive conditional heteroskedasticity (GARCH) models are used to exam the time-varying conditional variances. 

## Results
```{r, warning=FALSE}
load("timberland.Rdata")
library(tseries)
library(forecast)
library(dygraphs)
library(PerformanceAnalytics)
require(ggplot2)
require(GGally)
library(psych)
```

## 3.1 Descriptive analysis 
This is a interactively graphical representation of selected quarterly return indices from 1987 Q1 to 2020 Q1. Some clear pattern could be observed. First, NCREIF is much less volatile than REIT, SP500, RU2000. Second, REIT follows SP500 closely but NTI doesn't. For example, both REIT and SP500 suffered from a big loss during 2008 financial crisis, whereas NCREIF kept a positive returns until 2009/Q2, where they only got a slightly negative return value. Some people said it's because timberland market tends to lag the public equity market. But some also argues that it's because of the biological growth of trees wouldn't stop as the market went into depression. 
```{r}
dygraph(allxts*100, main = "Quarterly returns of analyzed assets", ylab = "Return%")%>% 
  dyHighlight(highlightSeriesOpts = list(strokeWidth = 3))%>%
  dyRangeSelector()%>%
  dyLegend(width = 500)%>%
  dyOptions(colors = RColorBrewer::brewer.pal(6, "Dark2"))

```

The unconditional correlations of the return series were studied by calculating the correlation coefficients and by plotting the pairwise scatter plots, as showed in the following graph. In this portfolio, the correlation of NCREIF with all financial assets are uniformly low, except for the 3 month Treasury bill (CA).  
```{r, message=FALSE}
ggpairs(as.data.frame(allts))+theme_bw()
```


```{r}
describe(allxts)[,-1]

VaR(allts, p=.95, method="modified")
ES(allts, p=.95, method="modified")
# normality
apply(allts,2,jarque.bera.test)
# stationarity
apply(allts,2,adf.test)
apply(allts,2,pp.test)


#test stationarity and normality
for(i in 1:ncol(allxts)){
  #plot acf
  acf(na.omit(allxts[,i])^2,lag.max = 100,xlab = colnames(allxts)[i], ylab = 'ACF', main=' ')
  #plot pacf
  pacf(na.omit(allxts[,i])^2,lag.max = 100,xlab = colnames(allxts)[i], ylab = 'PACF', main=' ')
}

```


## 3.2 Portfolio Optimization






